Emittance preservation of an electron beam in a loaded quasi-linear plasma wakefield

### 2017-09-03 Tex/Paper.tex
Added Erik's suggestions for parameter section and conclusion

In these 3D simulations we have used a matched beam for an initial emittance of $2\unit{\mu m}$, corresponding to a
$\sigma_{x}$ of $5.25\unit{\mu m}$. The beam matching relation is given by
\begin{align}
    \beta = \frac{\sigma_r^2}{\epsilon_g} = \frac{\lambda_pe}{2\pi}\sqrt{2\gamma_{rel}}, \label{EQ:Matched}
\end{align}
where $\lambda_{p}$ is the plasma wavelength. This is a very narrow beam compared to the drive beam
$\sigma_{x} = 200\unit{\mu m}$. The charge density of this compact witness beam reaches that of the plasma density at
only a few $\unit{pC}$, but reaches optimal beam loading at around $1-200\unit{pC}$. The implication here being that the
witness beam produces its own non-linear wakefield already at the head of the beam. The majority of the electrons
within it will therefore see a linear focusing force preventing the bulk of the beam from undergoing emittance growth.

@@

For accelerator applications in general, and also for Run 2 of the AWAKE experiment, it is desirable to maximise the
charge that can be successfully accelerated in the witness beam. However, a longitudinally constant accelerating
field depends on both the position and the charge density of the witness beam \cite{katsouleas:1987, tzoufras:2009}.

The large number of parameters involved makes the problem difficult to solve in simulations, and difficult to measure in
experiments. Presented in figures \ref{Fig:BeamQ} and \ref{Fig:BeamQAbs} is a parameter scan for a matched beam with
initial normalised emittance $\epsilon_{N,0}=2\unit{\mu m}$ after propagating through $4\unit{m}$ of plasma. We ran the
scan with beam charge from $10\unit{pC}$ to $300\unit{pC}$ and with length $\sigma_{z}$ from $40\unit{\mu m}$ to
$100\unit{\mu m}$. We see from figure \ref{Fig:BeamQ} that both the $40\unit{\mu m}$ and the $60\unit{\mu m}$ beam has a
well defined minimum energy spread with a beam charge $\geq 50\unit{pC}$ and $\approx 100\unit{pC}$ respectively. Lower
beam charges tend to underload the field, while higher beam charges tend to overload. It is also clear that long beams
with respect to the accelerating flank of the $E_{z}$-field, $\approx\lambda_{pe}/4$, will not optimally load the field
along its length thus experiencing a larger spread in energy.

A higher charge beam will generate a non-linear wakefield closer to the head of the beam, which in turns ensures that a
smaller portion of the beam experiences the quasi-linear region with emittance growth. As figure \ref{Fig:BeamQAbs}
illustrates, a higher charge also results in a lower energy gain.

@@

>>>>

In these 3D simulations we have used a matched beam for an initial emittance of $2\unit{\mu m}$, corresponding to a
$\sigma_{x}$ of $5.25\unit{\mu m}$. The beam matching relation is given by
\begin{align}
    \beta = \frac{\sigma_r^2}{\epsilon_g} = \frac{\lambda_{pe}}{2\pi}\sqrt{2\gamma_{rel}}, \label{EQ:Matched}
\end{align}
where $\lambda_{p}$ is the plasma wavelength. This is a very narrow beam compared to the drive beam
$\sigma_{x} = 200\unit{\mu m}$. The charge density of this compact witness beam reaches that of the plasma density at
only a few $\unit{pC}$, but reaches optimal beam loading at around $1-200\unit{pC}$. The implication here being that the
witness beam produces its own non-linear wakefield already at the head of the beam. The majority of the electrons
within it will therefore see a linear focusing force preventing the bulk of the beam from undergoing emittance growth.

@@

The beam loading and blow out properties of the electron beam depends on a large number of parameters: the longitudinal
profile, the transverse profile as well as the relative phasing of the proton and the electron beams.

We here study parameter optimization relevant for AWAKE Run 2, in order to understand the ideal injection parameters.
For AWAKE Run 2 we desire to maximize the energy gain, minimize the energy spread, maximize the charge to be accelerated
and minimize the emittance growth. In addition, the bunch length should be such that it is possible to generate and
transport by a compact electron injector [IPAC]. These are contradictory constraints. We  investigate the
interdependence of these parameters by simulation. We vary the electron bunch length, charge and initial emittance. We
quantify the resulting emittance growth, energy spread and and energy gain. The desired injection parameters will depend
on the application.

Figures \ref{Fig:BeamQ} and \ref{Fig:BeamQAbs} show the dependence on bunch length and bunch charge for an initial beam
emittance of $2\unit{\mu m}$ after propagating through $4\unit{m}$ of plasma. We ran the scan with beam charges from
$10\unit{pC}$ to $300\unit{pC}$ and with length $\sigma_{z}$ from $40\unit{\mu m}$ to $100\unit{\mu m}$. We see from
figure \ref{Fig:BeamQ} that both the $40\unit{\mu m}$ and the $60\unit{\mu m}$ beam has a well defined minimum energy
spread with a beam charge $\geq 50\unit{pC}$ and $\approx 100\unit{pC}$ respectively. Lower beam charges tend to
underload the $E_{z}$-field, while higher beam charges tend to overload it. It is also clear that long beams with
respect to the accelerating flank of the $E_{z}$-field, $\approx\lambda_{pe}/4$, will not optimally load the field along
its length thus experiencing a larger spread in energy.

[Discussion of new emittance scan figure(s) to be added.]

@@

We have studied a new plasma wakefield acceleration regime; acceleration of a strongly loaded electron beam in a quasi-
linear proton driven wake. When optimally loaded, the electron beam creates its own ion bubble, yielding emittance
preservation for a large parts of the electron beam.

These result indicates that emittance preservation for an on-axis injected electron beam, properly optimized, is
feasible. Parameter studies indicate that up to a few $100\unit{pC}$ might be accelerated, for bunches of length
$40-60\unit{\mu m}$. Such an electron bunch may be generated by a standard S-band RF gun.

As this study assumes a wake driven by a single, short proton beam, the studies for interesting cases for AWAKE should
be repeated by a full simulation of a self-modulated proton beam.
